Quantifying Spatially Resolved Hydration Thermodynamics Using Grid Inhomogeneous Solvation Theory [Article v\versionnumber]

### 2022-12-21 manuscript/manuscript.tex
Add formula to double counting section, small changes

Densities of typical water models can be found in the \software{cpptraj} manual.
We also provide the densities for water models commonly used with AMBER forcefields in table \ref{tab:ref_densities}. These densities were calculated from 1$\mu$s MD simulations. Pressure was kept constant at 1 bar with the Berendsen barostat and the water box was prepared with tleap to a size of 100$\cdot$100$\cdot$100\AA before pressure equilibration. The temperature was kept constant at 300 K with the Langevin thermostat.
\begin{table}[h]
	\caption{Reference density $\rho_0$ and mean water-water energy $E_{ww, norm}$ for various water models from NpT simulations at 300K/1bar.}\label{tab:ref_densities}
	\small
	\begin{tabularx}{\columnwidth}{@{}lrr@{}}
		\toprule
		Water Model       & $\rho_0$ /\AA$^{-3}$ & $E_{ww, norm}$ /$kcal\cdot mol^{-1} \cdot water^{-1}$  \\
		\midrule
		TIP3P    & 0.0329 & -9.533   \\
		TIP3PFB  & 0.0334 & -11.369  \\
		TIP4P    & 0.0332 & -9.856   \\
		TIP4P/Ew & 0.0332 & -11.036  \\
		TIP5P    & 0.0329 & -9.596   \\
		SPC/E    & 0.0333 & -11.123  \\
		OPC      & XXXX   & XXXX     \\
		OPC3     & XXXX   & XXXX     \\
		POL3     & XXXX   & XXXX     \\
		\bottomrule
	\end{tabularx}
\end{table}
If your simulation settings diverge significantly from the above setting or for other solvent models not listed either here or in the \software{cpptraj} manual, the density should be computed from a pure solvent simulation beforehand.

@@

\subsubsection{MD simulation}
\todo{Move simulation part from complex to here}
Now, you can run a restrained MD simulation based on the equilibrated structure of biotin for the subsequent GIST analysis.
Restraints should be used during the MD because IST theory is not defined for moving solutes.
We apply a harmonic restraint of \SI{100}{\kilo\calorie\per\mole\per\angstrom\squared} to all heavy atoms to keep the system in place.
We run \SI{100}{\nano\second} of simulation and store the conformations every \SI{100}{\pico\second}.
A simulation length of at least \SI{20}{\nano\second} is recommended to obtain statistically independent snapshots of the water molecules.

@@

To run PME-GIST, the \inlinecode{pme} flag to \inlinecode{gist} can be used in \software{cpptraj}.
The output will contain two additional columns: \inlinecode{PME\_dens} and \inlinecode{PME\_norm}.
They contain the potential of solute and solvent evaluated at the solvent positions and divided by two.
Furthermore, the \inlinecode{Eww} and \inlinecode{Esw} columns are also computed using PME, and are reported as usual.
For typical use cases, we recommend using \inlinecode{Eww} and \inlinecode{Esw} over the \inlinecode{PME} column.

\subsection{GIST with non-water solvents}
In References~\cite{Kraml2020,Kamenik2020-gist-macrocycles,Waibl2022-gist-solvents}, an extension of GIST towards solvents other than water is described.

@@

\begin{equation}
	\Delta E^\text{total}_{ww} = \sum_{k}^\text{voxels} \frac{E_{ww,k}^\text{corr}}{2}
%	\Delta E^R_{ww} = n^R E^\text{bulk}_{ww} - E^\textit{R,corr}_{ww}
\end{equation}

@@

\begin{equation}
	\Delta E^\text{total}_{ww} = \sum_{k}^{\text{voxels in }R} E_{ww,k}^\text{corr}
%	\Delta E^\textit{R,norm}_\textit{ww} = E^\textit{R,norm}_\textit{ww} - 2E^\textit{bulk}_\textit{ww}
\end{equation}

When the region $R$ comprises more than one molecule, interactions within this region are double-counted while interactions to the outside are not.
This could be solved using the solvent-solvent energy between each pair of voxels $E_{ww,kl}$.
While this is supported by the standard GIST implementation, it is rarely done due to the large size of the $E_{ww,kl}$ matrix.

>>>>

Densities of typical water models can be found in the \software{cpptraj} manual.
We also provide the densities for water models commonly used with AMBER forcefields in table \ref{tab:ref_densities}. These densities were calculated from 1$\mu$s MD simulations. Pressure was kept constant at 1 bar with the Berendsen barostat and the water box was prepared with tleap to a size of 100$\cdot$100$\cdot$100\AA\ before pressure equilibration. The temperature was kept constant at 300 K with the Langevin thermostat.
\todo{vah -> Finish these calculations}
\begin{table}[h]
	\caption{Reference density $\rho_0$ and mean water-water energy $E_{ww, norm}$ for various water models from NpT simulations at 300K/1bar.}\label{tab:ref_densities}
	\small
	\begin{tabularx}{\columnwidth}{@{}lrr@{}}
		\toprule
		Water Model       & $\rho_0$ /\AA$^{-3}$ & $E_{ww, norm}$ /$kcal\cdot mol^{-1} \cdot water^{-1}$  \\
		\midrule
		TIP3P    & 0.0329 & -9.533   \\
		TIP3PFB  & 0.0334 & -11.369  \\
		TIP4P    & 0.0332 & -9.856   \\
		TIP4P/Ew & 0.0332 & -11.036  \\
		TIP5P    & 0.0329 & -9.596   \\
		SPC/E    & 0.0333 & -11.123  \\
		OPC      & XXXX   & XXXX     \\
		OPC3     & XXXX   & XXXX     \\
		POL3     & XXXX   & XXXX     \\
		\bottomrule
	\end{tabularx}
\end{table}
If your simulation settings diverge significantly from the above setting or for other solvent models not listed either here or in the \software{cpptraj} manual, the density should be computed from a pure solvent simulation beforehand.

@@

\subsubsection{MD simulation}

Now, you can run a restrained MD simulation based on the equilibrated structure of biotin for the subsequent GIST analysis.
Restraints should be used during the MD because IST theory is not defined for moving solutes.
We apply a harmonic restraint of \SI{100}{\kilo\calorie\per\mole\per\angstrom\squared} to all heavy atoms to keep the system in place.
We run \SI{100}{\nano\second} of simulation and store the conformations every \SI{100}{\pico\second}.
A simulation length of at least \SI{20}{\nano\second} is recommended to obtain statistically independent snapshots of the water molecules.

@@

To run PME-GIST, the \inlinecode{pme} flag to \inlinecode{gist} can be used in \software{cpptraj}.
The output will contain two additional columns: \inlinecode{PME\_dens} and \inlinecode{PME\_norm}.
They contain the potential of solute and solvent evaluated at the solvent positions and divided by two.
Furthermore, the \inlinecode{Eww} and \inlinecode{Esw} columns are also computed using PME, and are reported as usual.
For typical use cases, we recommend using \inlinecode{Eww} and \inlinecode{Esw} over the \inlinecode{PME} column.
\todo{vah: The cpptraj manual says that GIST PME energies cannot be decomposed into Esw and Eww -> what's true? Seems to conflict with what is written in this section.}
\subsection{GIST with non-water solvents}
In References~\cite{Kraml2020,Kamenik2020-gist-macrocycles,Waibl2022-gist-solvents}, an extension of GIST towards solvents other than water is described.

@@

\begin{equation}
	\Delta E^\text{R}_{ww} = \sum_{k}^\text{voxels} \frac{E_{ww,k}^\text{corr}}{2}
%	\Delta E^R_{ww} = n^R E^\text{bulk}_{ww} - E^\textit{R,corr}_{ww}
\end{equation}

@@

\begin{equation}
	\Delta E^\text{R}_{ww} = \sum_{k}^{\text{voxels in }R} E_{ww,k}^\text{corr}
%	\Delta E^\textit{R,norm}_\textit{ww} = E^\textit{R,norm}_\textit{ww} - 2E^\textit{bulk}_\textit{ww}
\end{equation}

When the region $R$ comprises more than one molecule, interactions within this region are double-counted while interactions to the outside are not.
This could be solved using the solvent-solvent energy between each pair of voxels $E_{ww,k\;l}$:
\begin{equation}
	\Delta E^\text{R}_{ww} = 2\left(\sum_{k}^\text{voxels in R} E_{ww,k}^\text{corr}-\sum_{k}^\text{voxels in R}\sum_{l}^\text{l>k} E_{ww, k\;l}\right)
\end{equation}
While this is supported by the standard GIST implementation, it is rarely done due to the large size of the $E_{ww,k\;l}$ matrix.
